---
title: "Population synthesis on the basis of Gibbs-sampling"
output: html_notebook
---

```{r}
################ Setting up the data: Illustration of Gibbs-sampling   ################

install.packages("mipfp")
install.packages("plyr")
install.packages("plotrix")
install.packages("tidyr")
install.packages("ztable")

require("mipfp")
require("plyr")
require("plotrix")
require("tidyr")
require("ztable")
```


When creating a synthetic population of agents over some feature-set, it is possible to build a pool of agents (with these features) and then apply some sort of re-sampling from that pool. 

A way to approximate such distributions is by using Gibbs sampling, where we draw from margins. For a very simple 3-by-3 case it may look like. 


```{r}

##############################################
# Case1: Implementing a simple Gibbs-sampler #
##############################################

#Initialise data 

initial.guess <- c(53, 39, 60, 40, 78, 83, 29, 98, 80)
#initial.guess0 <- c(1, 1, 1, 1, 1, 1, 1, 1, 1)

dim(initial.guess) = c(3,3)
initial.guess = t(initial.guess)

df = data.frame(initial.guess) # Converting to dataframe for data modification

inc_cat = c('inc_1', 'inc_2', 'inc_3') # Names of income categories 
age_cat = c('age_1', 'age_2', 'age_3') # Names of age categories

colnames(df) <- age_cat # Change the names of the columns of the dataframe
df = cbind(inc_cat, df) # Add a column with income names

data_long <- gather(df, age_cat, values, age_1:age_3, factor_key=TRUE) # Change dataframe format from wide to long (three columns, two with variable names, one with values per combination)

data_long$inc_cat <- as.character(data_long$inc_cat) # Strings are easier to work with than factorial variables
data_long$age_cat <- as.character(data_long$age_cat) # Strings are easier to work with than factorial variables

num_samples = 1000 
sampled_individuals <- data.frame(matrix(NA, nrow = num_samples+1, ncol = 2)) # Initialize matrix
colnames(sampled_individuals) <- list('inc_cat', 'age_cat') # Change names

sampled_individuals[1,1] <- sample(data_long[,'inc_cat'], 1) # Initialize the gibbs sampler

print(sampled_individuals[1,1])

```

Now define the loop from which the drawing from the conditionals are carried out.

```{r}
# Loop when the conditional sampling is carried out
for (n in 1:num_samples){
  
## Draw for age 
  condition_on_inc = data_long[which(data_long[,'inc_cat']==sampled_individuals[n, 'inc_cat']),]
  sampled_individuals[n, 'age_cat'] = sample(condition_on_inc[,'age_cat'], 1, prob = condition_on_inc$values/sum(condition_on_inc$values))
  
## Draw for inc 
  condition_on_age = data_long[which(data_long[,'age_cat']==sampled_individuals[n, 'age_cat']),]
  sampled_individuals[n+1, 'inc_cat'] = sample(condition_on_age[,'inc_cat'], 1, prob = condition_on_age$values/sum(condition_on_age$values))
}

# Storing the 1000 individuals to a dataframe

Gibbs_sample = sampled_individuals[1:1000,]
head(Gibbs_sample,10)

#Adding an ID
Gibbs_sample$ID <- seq.int(nrow(Gibbs_sample))

Gibbs_table <- table(Gibbs_sample$inc_cat,Gibbs_sample$age_cat)

# Use Ztable for nice heatmap presentation
ztable(Gibbs_table) %>%
  makeHeatmap() %>% print(caption="Gibbs heatmap")

# Print with caption
print(Gibbs_table,caption="Gibbs sampled frequencies")

#### Similar for Initial! ####

# Different table functions
margin.table(Gibbs_table, 1) # A frequencies (summed over B)
margin.table(Gibbs_table, 2) # B frequencies (summed over A)

prop.table(Gibbs_table) # cell percentages

prop.table(Gibbs_table, 1) # row percentages
prop.table(Gibbs_table, 2) # column percentages

```

Now what to do with this pool of people? Its essentially just reflecting the distribution of the starting solution and the rendering of an agent-based representation!

The trick is now to apply re-sampling from that pool

```{r}
#Set starting solution

initial.guess <- c(53, 39, 60, 40, 78, 83, 29, 98, 80)
#initial.guess0 <- c(1, 1, 1, 1, 1, 1, 1, 1, 1)

dim(initial.guess) = c(3,3)
initial.guess = t(initial.guess)

#Set margins 
target.income <- c(3000, 3200, 3800)
target.age <- c(2000, 4000, 4000)

target.data.2005 <- list(target.income, target.age)

print(target.data.2005)

#Set the dimensions of the target data, income is first and age is second
target.dim <- list(1,2)

#Call the IPF
Fit_3_3_2005 <- Ipfp(initial.guess, target.dim, target.data.2005, print = TRUE, tol = 0.0011)

#Dump IPF fitted array
Fit <- Fit_3_3_2005$x.hat
Prob <- Fit_3_3_2005$p.hat

#Convert to Data frame
FitDB <- adply(Fit, c(1,2), id = NA)
ProbDB <- adply(Prob, c(1,2), id = NA)

#Rename columns
ProbDB <- rename(ProbDB, c("X1" = "income", "X2"="age", "V1"="Prob"))

# First create the new field
ProbDB$inc_cat <- NA
ProbDB$age_cat <- NA

# Then recode the old field into the new one for the specified rows
ProbDB$inc_cat[ProbDB$income==1] <- "inc_1"
ProbDB$inc_cat[ProbDB$income==2] <- "inc_2"
ProbDB$inc_cat[ProbDB$income==3] <- "inc_3"

ProbDB$age_cat[ProbDB$age==1] <- "age_1"
ProbDB$age_cat[ProbDB$age==2] <- "age_2"
ProbDB$age_cat[ProbDB$age==3] <- "age_3"

#Merge IPF weights/probabilities to the original data
Gibbs_sample <- merge(Gibbs_sample, ProbDB, by = c("inc_cat", "age_cat"), all.x=1, sort=1)

#Re-sampling with replacement
set.seed(321)
Gibbs_sample1000 <- sample(Gibbs_sample$ID,replace = TRUE, prob = Gibbs_sample$Prob, 1000)

#Convert to Data frame
Gibbs_sample1000 <- adply(Gibbs_sample1000, c(1), id = NA)

#Rename columns
names(Gibbs_sample1000)[1:2] <- c("ID_samp", "ID")

#Merge to initial structure: the sample is now a weighted sample with IPF weights
Gibbs_sample1000 <- merge(Gibbs_sample1000, Gibbs_sample, by = c("ID"), all.x=0, sort=1)

head(Gibbs_sample1000, 10)

#Analyse the structure of the population

```

