---
title: "Iterative proportional fitting for population syntheis: Week 2"
output: html_notebook
---

This is a R Markdown Notebook for the illustration of how IPF can be used to fit multi-dimensional matrices to be used for population synthesis. 

```{r}
################ Load and require packages ################

install.packages("mipfp")
install.packages("plyr")
install.packages("plotrix")
install.packages("data.table")

```

```{r}
require("mipfp")
require("plyr")
require("dplyr")
require("plotrix")
require("data.table")
library(openxlsx) # Library package statement

```

Define the work-directory

```{r setup}
knitr::opts_knit$set(root.dir = 'C:/Users/Popof/Desktop/course notes/Week2 - Population synthesis')
```

As a first example we will try to fit a 3-by-3 matrix according to row and column constrains

```{r}
###########################################
# Case1: Simple small 3-by-3 case example #
###########################################

#Set starting solution

initial.guess <- c(53, 39, 60, 40, 78, 83, 29, 98, 80)
#initial.guess <- c(1, 1, 1, 1, 1, 1, 1, 1, 1)

dim(initial.guess) = c(3,3)
initial.guess = t(initial.guess)

#Set margins 
target.income <- c(3000, 3200, 3800)
target.age <- c(2000, 4000, 4000)

target.data.2005 <- list(target.income, target.age)

print(target.data.2005)

#Set the dimensions of the target data, income is first and age is second
target.dim <- list(1,2)

```

Now we have organised the data for fitting the matrix!

```{r}
#Call the IPF
Fit_3_3_2005 <- Ipfp(initial.guess, target.dim, target.data.2005, print = TRUE, tol = 0.0011)

```

The algorithm converged!Now lets get the data into a proper form.

```{r}

#Dump IPF fitted array
Fit <- Fit_3_3_2005$x.hat

#Reguire package
require("plyr")

#Convert to Data frame
Fit_3_3 <- adply(Fit, c(1,2), id = NA)

#Rename columns
names(Fit_3_3) <- c("Income", "Age", "FitVal")

head(Fit_3_3, n=10)

```

Now it is time to introduce a large-scale IPF case for Copenhagen. First we arrange data.

```{r}

##########################################
# Case4: large-scale example             #
##########################################

#Lib reference

#Importing csv
init <- read.csv(file="C:/Users/Popof/Desktop/course notes/Week2 - Population synthesis/Data - Handout 1/Init_Matrix_imputed.csv", header=TRUE, sep=",")

agezone17 <- read.csv(file="C:/Users/Popof/Desktop/course notes/Week2 - Population synthesis/Data - Handout 1/Target2017_AgeGender.csv", header=TRUE, sep=",")
incomezone17 <- read.csv(file="C:/Users/Popof/Desktop/course notes/Week2 - Population synthesis/Data - Handout 1/Target2017_AgeIncome.csv", header=TRUE, sep=",")

#Look at the files
head(init, n=10)
head(agezone17, n=10)
head(incomezone17, n=10)

#Define vectors
tgt.init <- init$Val
tgt.age <- agezone17$Val
tgt.income <- incomezone17$Val

dim(tgt.age) = c(32,2,10)
dim(tgt.income) = c(32,10,11)

# Organise with fastest moving index first
dim(tgt.init) = c(32,2,10,11)

tgt.list.dims <- list(c(1,2,3), c(1,3,4))
tgt.data <- list(tgt.age, tgt.income)

```

Then fit the model as before

```{r}
#Fit problem
Fit <- Ipfp(tgt.init, target.list=tgt.list.dims, target.data=tgt.data, print = TRUE, tol = 0.001, tol.margins = 0.001)

```

Now arrange the data 

```{r}
#Dump IPF fitted array
Fit17 <- Fit$x.hat
Prob17 <- Fit$p.hat

#Convert to Data frame
Fit17 <- adply(Fit17, c(1,2,3,4), id = NA)
Prob17 <- adply(Prob17, c(1,2,3,4), id = NA)

#Rename columns
names(Fit17)[1:5] <- c("ZID", "Male_Index", "AgeID", "Income_Index", "FitVal")
names(Prob17)[1:5] <- c("ZID", "Male_Index", "AgeID", "Income_Index", "ProbVal")

#Create a group index: The data keys does not conform for Male and IncomeID.
init$Male_Index <- group_indices(init, Male)
init$Income_Index <- group_indices(init, IncomeID)

#Merge to initial structure
Fit17 <- merge(init, Fit17, by = c("ZID", "Male_Index", "AgeID", "Income_Index"), all.x=0, sort=1)

#Sort
attach(Fit17)
Fit17 <- Fit17 [order(ZID, Male_Index, AgeID, Income_Index),]

# Print
head(Fit17, n=10)
detach(Fit17) 
```

Possibly export the data to external file

```{r}
#Possible export

#Excel
write.xlsx(Fit17, "./Pop synthesis/Population synthesis Case4/Results/Fit17.xlsx", sheetName="IPF17", col.names=TRUE)

#Export csv
#write.csv(Fit17, "./Pop synthesis/Population synthesis Case4/results/Fit17.csv")

#Fast export of CSV
fwrite(Fit17,"./Pop synthesis/Population synthesis Case4/results/Fit17.csv")

```


Until now we have only created IPF probabilities for a 2017 case. However, when synthezising and simulating population agents for the future, we will need to apply future targets.


```{r}

#Re-run IP for 2040...

```
The development of scenarios/targets is often not straightforward. As an example, how would we project a general income-growth when income are categorical and represent prototypes?

In that case we may apply a simple simulation trick!

```{r}
########################################################################################
####### Shift Income by 10% via simulation ########

### 10% increase in salary

h <- weighted.hist(Fit17$Income_Index, Fit17$Val,breaks=10,col=NULL,
                   plot=FALSE, freq=FALSE,ylim=NA,ylab=NULL,xaxis=TRUE)
barplot(c(h$density,0.),space=0)

xtick <- c(1,2,3,4,5,6,7,8,9,10,11)

text(x=xtick,  par("usr")[3], labels = xtick, pos = 1, xpd = TRUE)

# in this case we directly sample the population and increase their income by ten percent
j = 0
distrib = rep(0,11)

population_income <- list()

for (counts in h$counts){
  print(count)
  
  new_income <- runif(counts, j*100, (j+1)*100)
  population_income[[j+1]] <- new_income
  new_income <- new_income*1.1
  
  for (i in 1:length(distrib)){
    distrib[i] <- distrib[i] + sum( ((i-1)*100) < new_income & new_income<= (i*100))
  }
  
  j = j+1
}

lines(x = c(h$mids-1, 10.5), y = distrib/sum(distrib)) 
points(x = c(h$mids-1, 10.5), y = distrib/sum(distrib)) 

```



